---
title: "Course Project 2"
author: "Stefan Buchholz"
date: "Wednesday, October 08, 2014"
output: html_document
---

Questions to answer:

## Questions
***
You must address the following questions and tasks in your exploratory analysis. For each question/task you will need to make a single plot. Unless specified, you can use any plotting system in R to make your plot.

1. Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? Using the base plotting system, make a plot showing the total PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.

2. Have total emissions from PM2.5 decreased in the Baltimore City, Maryland (fips == "24510") from 1999 to 2008? Use the base plotting system to make a plot answering this question.

3. Of the four types of sources indicated by the type (point, nonpoint, onroad, nonroad) variable, which of these four sources have seen decreases in emissions from 1999–2008 for Baltimore City? Which have seen increases in emissions from 1999–2008? Use the ggplot2 plotting system to make a plot answer this question.

4. Across the United States, how have emissions from coal combustion-related sources changed from 1999–2008?

5. How have emissions from motor vehicle sources changed from 1999–2008 in Baltimore City?

6. Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in Los Angeles County, California (fips == "06037"). Which city has seen greater changes over time in motor vehicle emissions?

## Prequesits
***
Download, unzip and load the data from [Coursera-Site](https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2FNEI_data.zip "Data")

```{r 1. Downloading the stuff, results="hide"}
url = "https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2FNEI_data.zip"
#download.file(url, dest="exdata.zip")
unzip("exdata.zip", exdir="inputfiles", list=FALSE, overwrite=TRUE, unzip="internal")
NEI <- readRDS("inputfiles/summarySCC_PM25.rds")
SCC <- readRDS("inputfiles/Source_Classification_Code.rds")
```

## Answering Questions
***

1. Total PM2.5 from 1999-2008 in the US with base plotting system.
```{r Answering question 1}
#SUM of PM2.5 aggregating by year
test <- aggregate(Emissions ~ year, data = NEI, FUN = sum)
plot(test, 
     type="l",
     main = "Total PM2.5 Emissions by Year",
     xlab = "Year",
     ylab = "P2.5 - Emissions (EINHEIT)")
```

2. Have total Emissions decreased for the city of Baltimore in Maryland from 99-08?
```{r Answering question 2}
bm <- with(NEI, fips == "24510")
bm <- NEI[bm, ]
bm_data <- aggregate(Emissions ~ year, data = bm, FUN = sum)
plot(bm_data, 
     type="l",
     main = "Total PM2.5 Emissions by Year for Baltimore",
     xlab = "Year",
     ylab = "P2.5 - Emissions (EINHEIT)")
```

3. How do the sources increase or decrease for Baltimore
```{r Answering question 3}
bm_data <- aggregate(x=bm$Emissions, by = list(bm$year ,bm$type) ,FUN = sum)
colnames(bm_data) <- c("Year", "Source_Type", "Emission")
library(ggplot2)
gp <- ggplot(data = bm_data, aes(x=Year, y=Emission, group=Source_Type, color=Source_Type)) +geom_line() +geom_point()
gp <- gp + ggtitle("Total Emmissions by Pollution Source and Year for Baltimore")
```

4. How did coal combustion-related sources changed over time
```{r Answering question 4}
library(dplyr)
#just get the SCC column of the descriptions where One = External Combustion ... and coal in Level Three is used
finalfilter <- filter(SCC, SCC.Level.One == "External Combustion Boilers", SCC.Level.Three %in% grep("coal",ignore.case=T, value=T, SCC$SCC.Level.Three))
NEI_coal_combusted_filtered <- NEI[(NEI$SCC %in% finalfilter$SCC),]

# group by year
coalcombustedAggregated <- aggregate(Emissions ~ year, data = NEI_coal_combusted_filtered, FUN = sum)

# plot
library(ggplot2)
#png("plot4.png")
ggplot(coalcombustedAggregated, aes(x=year, y=Emissions)) + geom_line() +
    ggtitle("Emissions from coal combustion-related sources")
dev.off()
```

5. How did Motor Vehicle Sources in Baltimore over the time period behave?